Influence of Connectivity on Dung Beetle Communities

### 2025-04-09 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
quick question
update

Paragraph 2: What we know. 
  Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003].

@@

>Topic Sentence:  Changes in ecosystem fuction as a result of shifts in community structure

Paragraph 4: Why we don't know it. _(hint: ecosystem services)_

>Topic Sentence: Is this where we talk about challenges in large scale experimental design and the corridor project or is this where we talk about ecosystem services and biodiversity?

@@

Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al. 2024). The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]
Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes. We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge. To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors. Additionally we asked (2) Are corridors benefiting any one functional trait over another? Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling.

@@

All dung beetles were counted and identified to species with the exception of beetles of the genus Aphodius, overall 15 species were identified and approximately 5300 individual beetles were collected. 
 - what guides/keys did you do to ID? 
 - did you deposit vouchers and if so where are they available?

@@

1. how many beetles total
1. most common species overall. 
1. was. the same species the most common one in all habitats?
1. were there any that were caught in all habitats? Any restricted to only one?
1. any invasive / exotic species?
1. Table: count by species by habitat

@@

@@

>>>>

Paragraph 2: What we know.

Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003], including decreases in the abundance and diversity of species.

@@

<!-- stuff that these inviduals or species do: -->
<!-- 1) their interactions with other species -->
<!-- - predation -->
<!-- - pollination -->
<!-- - seed dispersal -->
<!-- 2) stuff on their own that influences ecosystems -->
<!-- - ecosystem engineering (beavers, leaf-cutter ants) -->
<!-- - nutrient cycling -->
<!-- - decomposition -->

>Topic Sentence:  Changes in ecosystem fuction as a result of shifts in community structure (EB but we may not want to emphasize this given we didn't measure directly)

Potential alternative routes:

1) We know far less on how corridors influence functional diversity. Functional diversity is important for the maintenance of ecosystems services, etc, so corridors cvould help ensure that.  (in context of dung beetles that would be feeding guilds (simplest), morphology etc. in deCastro Arrazola et al)

2) Studies investigating corridors rarely look at whats i the matrix. The matrix could be a source of diversity that moves into patches and then through corridors or a completely different subset of species. (any other options?) Understanding whichh of these it is is important because it leads to differnt conclusions abiout role of corridors and conservation strategies

3) regardless, not a bad idea to discuss both matrix and functional group in diuscussion.

4) and the alternative approach: revamp into around spillover from matrix into patches and if depednent on corridors or edge

<!-- Paragraph 4: Why we don't know it. _(hint: ecosystem services)_   -->

<!-- >Topic Sentence: Is this where we talk about challenges in large scale experimental design and the corridor project or is this where we talk about ecosystem services and biodiversity?  -->

@@

Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al. 2024). The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]
Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes. We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge. To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors. Additionally we asked (2) Are corridors benefiting any one functional trait over another? 
<!-- Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling. -->

@@

All dung beetles were counted and identified to species. 
 - We also dried and weighed individuals to estimate biomass 
 - what guides/keys did you do to ID? 
 - did you deposit vouchers and if so where are they available?

@@

### Overview
_Don't have to describe these, just do it and report in results. this is just to remind us of what to do_
1. how many beetles total
1. most common species overall. 
1. was. the same species the most common one in all habitats?
1. were there any that were caught in all habitats? Any restricted to only one?
1. any invasive / exotic species?
1. Table: count by species by habitat

@@

To Test for differences in species richness, we....

@@

We used a similar approach to compare species diversity in winged, isolated, and connected patches....
